---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ameras

<!-- badges: start -->
[![R-CMD-check](https://github.com/sanderroberti/ameras/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sanderroberti/ameras/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of ameras is to provide a user-friendly interface to analyze association studies with multiple replicates of a noisy exposure using a variety of methods. ameras supports continuous, count, binary, multinomial, and right-censored time-to-event outcomes. For binary outcomes, the nested case-control design is also accommodated. Besides the common exponential relative risk model $RR=\exp(\beta D)$ for the exposure-outcome association with noisy exposure $D$, linear excess relative risk $RR=1+\beta D$ and linear-exponential excess relative risk models $RR=1+\beta_1 D \exp(\beta_2 D)$ can be used. 



## Installation

To install from CRAN:
```{r install}
# install.packages("ameras")
```


## Example

This is a basic example which shows you how to fit a simple logistic regression model:

```{r example}
library(ameras)
data(data, package="ameras")
dosevars <- paste0("V", 1:10)
fit <- ameras(data, family="binomial", Y="Y.binomial", methods=c("RC","ERC","MCML", "FMA", "BMA"), 
              dosevars=dosevars)
summary(fit)
```
