---
title: "Confidence intervals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Confidence intervals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ameras)

```

# Introduction
To compute confidence intervals, first fit the model using `ameras`, then use the `confint` method to attach confidence intervals.
Several types of confidence intervals are supported, which should be supplied to the `type` argument of `confint`, see below. When `confint` is called with `methods` containing at least one of `RC`, `ERC`, and `MCML` and at least one of `FMA` and `BMA`, `type` should be a vector of length 2 with one method for `RC`, `ERC` and `MCML` and one for `FMA` and `BMA`.

# Regression calibration, extended regression calibration, and Monte Carlo maximum likelihood

For (extended) regression calibration and Monte Carlo maximum likelihood, there are two types of Wald intervals, obtained either before or after transformation. If no transformation is specified, `wald.orig` should be used to obtain the standard Wald intervals. When a transformation is used, `wald.transformed` is determined before transforming, and `wald.orig` is obtained after transforming using the delta method (using `transform.jacobian`). The third option is `proflik`, which uses the profile likelihood to compute confidence bounds. For this, the profile log (partial) likelihood for parameter $\theta_p$ is defined as
$$
    PL_p (\theta_p^*) = \max_{\mathbf\theta: \theta_p = \theta_p^*} \ell (\mathbf \theta),
$$
where $\ell$ is the log (partial) likelihood. Next, profile confidence intervals $(\theta_p^l, \theta_p^h)$ are obtained for parameter $\theta_p$ at significance level $\alpha=0.95$ by solving 
$-2 \{PL_p(\theta_p^*) - \ell(\hat{\mathbf{\theta}})\}=\chi^2_{1,1-\alpha}$ using the bisection method, with $\hat{\mathbf{\theta}}$ the maximum likelihood estimate. Note that profile likelihoods are more computationally intensive to obtain. For this reason, `confint` offers the option to only determine them for the exposure-related parameters, which is the default setting. To obtain profile likelihood intervals for all parameters, use `parm = "all"`.

To illustrate, we determine the three types of confidence intervals for a regression calibration analysis using the example data, using a linear excess relative risk model with the default exponential transformation (see [Parameter transformations](transformations.html)).

```{r fits1, eval = identical(Sys.getenv("NOT_CRAN"), "true")}
data(data, package="ameras")
dosevars <- paste0("V", 1:10)
fit.ameras <- ameras(Y.binomial~dose(V1:V10, model="ERR")+X1+X2, data=data, 
                            family="binomial", methods=c("RC"))

fit.ameras.waldorig <- confint(fit.ameras, type="wald.orig")
fit.ameras.waldtransformed <- confint(fit.ameras, type="wald.transformed")
fit.ameras.proflik <- confint(fit.ameras, type="proflik", parm="all")

summary(fit.ameras.waldorig)
summary(fit.ameras.waldtransformed)
summary(fit.ameras.proflik)
```
# Frequentist and Bayesian model averaging

For frequentist and Bayesian model averaging methods, the options are `percentile` which uses 2.5\% and 97.5\% percentiles, and `hpd` which computes highest posterior density intervals using `HPDinterval` from the `coda` package, using either the FMA samples or Bayesian posterior samples.

Again, we use the example data to illustrate.

```{r fits2, eval = identical(Sys.getenv("NOT_CRAN"), "true")}
fit.ameras2 <- ameras(Y.binomial~dose(V1:V10, model="ERR")+X1+X2, data=data, 
                            family="binomial", methods=c("FMA"))

fit.ameras.hpd <- confint(fit.ameras2, type="hpd")
fit.ameras.percentile <- confint(fit.ameras2, type="percentile")

summary(fit.ameras.hpd)
summary(fit.ameras.percentile)
```